---
title: "Taylor Swift Album Cover Comparisons"
author: "Andrew Lucas"
---

Comparing color composition, similarities, and differences across all 15 Taylor Swift album covers in the dataset.

## Color Category Distribution

Rather than measuring isolated hex codes, colors are grouped into broad families (neutrals, warm tones, and accents).

```{r}
#| message: false
#| warning: false
#| echo: false
library(tidyverse)
```

```{r}
#| code-fold: true
#| code-summary: "Load the color data and declare specific color mappings"
# Load categorized color percentages for all 15 covers
album_colors <- read_csv("data/album_color_summary.csv", show_col_types = FALSE) |>
  pivot_longer(cols = -album, names_to = "color_group", values_to = "percentage") |>
  filter(percentage >= 3) # Keep prominent color categories

# Map each category name directly to its representative color
category_colors <- c(
  "Black"            = "#1a1a1a",
  "Gray"             = "#8a8a8e",
  "White"            = "#f5f5f7",
  "Beige / Cream"    = "#f3e9d2",
  "Brown"            = "#582f0e",
  "Brown / Tan"      = "#a68a64",
  "Red"              = "#b7092a",
  "Orange / Gold"    = "#e09f3e",
  "Yellow"           = "#ffd166",
  "Green"            = "#2d6a4f",
  "Blue"             = "#3a86ff",
  "Purple"           = "#7209b7",
  "Pink"             = "#ff70a6",
  "Pastel / Neutral" = "#e2cfea",
  "Muted Slate"      = "#64748b"
)
```

```{r}
ggplot(album_colors, aes(x = percentage, y = fct_reorder(album, percentage, sum), fill = color_group)) +
  geom_col(position = "fill", color = "grey50", linewidth = 0.2) +
  scale_fill_manual(values = category_colors) +
  scale_x_continuous(labels = scales::percent) +
  labs(
    title = "Relative Color Composition Across Album Covers",
    x = "Proportion",
    y = NULL,
    fill = "Color Category"
  ) +
  theme_minimal()
```

## Similarities and Differences

- **Monochrome & High Contrast:** *folklore*, *reputation*, and *THE TORTURED POETS DEPARTMENT* are closely linked, each consisting of over 85% grayscale tones (Black, White, Gray).
- **Warm & Earthy:** *Fearless*, *Fearless (Taylor's Version)*, *evermore*, and *Taylor Swift* share heavy concentrations of Gold, Brown, and Beige/Cream.
- **Vibrant Accents:** *Speak Now* and *Speak Now (Taylor's Version)* stand out with signature Purple and Pink tones, while *Red (Taylor's Version)* introduces deep red accents against dark neutrals.
- **Original vs. Taylor's Version:** The re-recorded covers systematically shift toward warmer, darker, or more muted palettes compared to their higher-key originals (e.g., *Fearless TV* deepening from light gold/white to rich brown and gold).

